---
title: "OnDeck Exercise"
output: html_document
---
Loading in the data--easier to read and manipulate with plyr

```{r}
data <- read.csv("OnDeck Analytics Asssignment.csv")
library(dplyr)
plydata <- tbl_df(data) 
head(plydata)
summary(plydata)
```

Cut delinquent days into factors, inclusive on the right side

```{r}
plydata$days_delinquent_old <- cut(plydata$days_delinquent_old, c(0,1,5,10,30,60,180), labels = c("0", "1-5", "5-10", "10-30", "30-60", "60+"), right=FALSE, include.lowest=TRUE) 
plydata$days_delinquent_new <- cut(plydata$days_delinquent_new, c(0,1,5,10,30,60,180), labels = c("0", "1-5", "5-10", "10-30", "30-60", "60+"), right=FALSE, include.lowest=TRUE)
```

Creating a transition matrix--note NaNs because there were no 0 day delinquencies to begin with

```{r}
change <- select(plydata, days_delinquent_new, days_delinquent_old)
library(markovchain)
mcX<-markovchainFit(change)$estimate
mcX
```

Create another transition matrix showing the probability of movement from one group to another, weighted by outstanding principal balance.

This should really be the old outstanding principal because we're looking at a time series to predict loan behavior

```{r}
principal <- plydata$new_outstanding_principal_balance
second <- cbind(change, principal)
```

Standardizing principal to percent of mean

```{r}
second$principal <- second$principal/mean(second$principal)
```

Discrete values so you can estimate the transition probabilities by the sample proportions

```{r}
library(questionr)
weightcount <- wtd.table(second$days_delinquent_old, second$days_delinquent_new, weights=second$principal)
weightcount <- weightcount/rowSums(weightcount)   
weightcount
```

Would expect the coverage ratio (approximated here by avg bank balance/new outstanding principal) to be key in determining loan performance

```{r}
plydata$coverage <- plydata$average_bank_balance__c/plydata$new_outstanding_principal_balance
```

Subsetting the data to begin investigating those loans that go from 10-30 days overdue to to 30-60 days

```{r} 
test <- subset(plydata, days_delinquent_old=="10-30")
test$worse <- test$days_delinquent_new == "30-60"
```

Modeling fico here as well

```{r}
fit <- lm(formula = worse ~ fico + coverage, data = test)
summary(fit)
```

Coverage is statistically significant (but not fico)--worth further investigation